---
title: "Data624 Assignment8"
author: "Ritesh Lohiya"
date: "April 12, 2019"
output: html_document
---

#Data624 Assignment 8


```{r}
suppressMessages(suppressWarnings(library(fpp2)))
suppressMessages(suppressWarnings(library(readxl)))
suppressMessages(suppressWarnings(library(seasonal)))
suppressMessages(suppressWarnings(library(rdatamarket)))
suppressMessages(suppressWarnings(library(tseries)))
suppressMessages(suppressWarnings(library(AppliedPredictiveModeling)))
suppressMessages(suppressWarnings(library(fma)))
suppressMessages(suppressWarnings(library(corrplot)))
suppressMessages(suppressWarnings(library(caret)))
suppressMessages(suppressWarnings(library(pls)))
suppressMessages(suppressWarnings(library(glmnet)))
suppressMessages(suppressWarnings(library(missForest)))
suppressMessages(suppressWarnings(library(mlbench)))
suppressMessages(suppressWarnings(library(nnet)))
suppressMessages(suppressWarnings(library(earth)))
suppressMessages(suppressWarnings(library(kernlab)))
suppressMessages(suppressWarnings(library(MASS)))
suppressMessages(suppressWarnings(library(forecast)))
suppressMessages(suppressWarnings(library(tidyverse)))
suppressMessages(suppressWarnings(library(Metrics)))
suppressMessages(suppressWarnings(library(Amelia)))
suppressMessages(suppressWarnings(library(doParallel)))
suppressMessages(suppressWarnings(library(cwhmisc)))
```

####7.2 Friedman (1991) introduced several benchmark data sets created by simulation. One of these simulations used the following nonlinear equation to create data:



$$y = 10sin(\pi x_1 x_2) + 20(x_3 -0.5)^2+10x_4 +5x_5 + N(0,\sigma^2)$$

####where the ex values are random variables uniformly distributed between [0,1] (there are also 5 0other non-informative variables also created in simulation). The package mlbench contains a function called mlbench.friedman1 that simulates these data:

```{r}
set.seed(200)
trainingData <- mlbench.friedman1(200, sd = 1)
## We convert the 'x' data from a matrix to a data frame
## One reason is that this will give the columns names.
trainingData$x <- data.frame(trainingData$x)
## Look at the data using
#featurePlot(trainingData$x, trainingData$y)
## or other methods.
```


```{r}
## This creates a list with a vector 'y' and a matrix
## of predictors 'x'. Also simulate a large test set to
## estimate the true error rate with good precision:
testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x)
```

#### KNN 

```{r}
knnModel <- train(x = trainingData$x, y = trainingData$y, method = "knn", preProc = c("center", "scale"), tuneLength = 10)
knnModel
```

####Make the prediction

```{r}
knnPred <- predict(knnModel, newdata = testData$x)
## The function 'postResample' can be used to get the test set
## perforamnce values
postResample(pred = knnPred, obs = testData$y)
```

####The RMSE is 3.17

####MARS

```{r}
set.seed(123)  
mars_grid = expand.grid(degree =1:2, nprune=seq(2,14,by=2))
marsTune = train(x= trainingData$x, y=trainingData$y, method='earth', tuneGrid=mars_grid, trControl = trainControl(method = "cv"))
marsTune
```

####Variable importance

```{r}
varImp(marsTune)
```


####Make the prediction using MARS.

```{r}
mars_pred = predict (marsTune, testData$x)
postResample(pred = mars_pred, obs = testData$y)
```

####The RMSE is 1.17, better than KNN

#### SVM


```{r}
set.seed(124)
svmTune = train(x= trainingData$x, y=trainingData$y, method='svmRadial', tuneLength = 14, trControl = trainControl(method = "cv"))
svmTune$finalModel
```

####Variable importance

```{r}
varImp(svmTune)
```


```{r}
svmPred <- predict(svmTune, newdata = testData$x)
postResample(pred = svmPred, obs = testData$y)
```

####The RMSE is 2.07

#### Which models appear to give the best performance? Does MARS select theinformative predictors (those named X1-X5)?

####With the above results, MARS is the most accurate of the 3 models. It has the lowest RMSE of 1.17.

#### 7.5.  Exercise 6.3 describes data for a chemical manufacturing process. Use the same data imputation, data splitting, and pre-processing steps as before and train several nonlinear regression models.

```{r}
data(ChemicalManufacturingProcess)
df = ChemicalManufacturingProcess
#summary(df)
```

```{r}
df_imp1 = missForest(df)
df_imp = df_imp1$ximp
data = df_imp[,2:58]
target = df_imp[,1]
training = createDataPartition( target, p=0.75 )
predictor_training = data[training$Resample1,]
target_training = target[training$Resample]
predictor_testing = data[-training$Resample1,]
target_testing = target[-training$Resample1]
```

####KNN

```{r}
knnModel1 <- train(x = predictor_training, y = target_training, method = "knn", preProc = c("center", "scale"), tuneLength = 10)
knnModel1
```

####The best RMSE is when n = 15.  


```{r}
knnPred1 <- predict(knnModel1, newdata = predictor_testing)
postResample(pred = knnPred1, obs = target_testing)
```

####The RMSE of 1.42.

#### MARS


```{r}
mars1 = expand.grid(degree =1:2, nprune=seq(2,14,by=2))
mars2 = train(x = predictor_training, y = target_training, method='earth', tuneGrid=mars1, trControl = trainControl(method = "cv"))
mars2
```

####Variable importance

```{r}
marspred = predict (mars2, predictor_testing)
postResample(pred = marspred, obs = target_testing)
```

#### The RMSE is 1.27, better than KNN model.

####SVM

```{r}
svm1 = train(x = predictor_training, y = target_training, method='svmRadial', tuneLength = 14, trControl = trainControl(method = "cv"))
svm1
```


```{r}
svmPred <- predict(svm1, newdata = predictor_testing)
postResample(pred = svmPred, obs = target_testing)
```

####The RMSE is 1.07 which is better then the previous models.

#### A. Which nonlinear regression model gives the optimal resampling and test set performance?

####The SVM model has the best accuracy with RMSE of 1.07


#### B. Which predictors are most important in the optimal nonlinear regression model? Do either the biological or process variables dominate the list? How do the top ten important predictors compare to the top ten predictors from the optimal linear model?

####Lets see the predictor importance of SVM model as this has the best RSME scores.

#### SVM top predictors

```{r}
plot(varImp(svm1))
v_imp <- varImp(svm1)
v_imp
```


####ManufacturingProcess32 and BiologicalMaterial06 are the top predictors. Both ManufacturingProcess and BiologicalMaterial contribute to the list if we consider the SVM algorithm.


#### C. Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model. Do these plots reveal intuition about the biological or process predictors and their relationship with yield?


```{r}
plot(data$ManufacturingProcess32, target)
abline(lm(target~data$ManufacturingProcess32),col="red",lwd=1.5)
```


```{r}

plot(data$BiologicalMaterial06, target)
abline(lm(target~data$BiologicalMaterial06),col="red",lwd=1.5)
```

Both predictors have positive corelation. Lets check the exact corelation

```{r}
cor(data$ManufacturingProcess32, target)
```

```{r}
cor(data$BiologicalMaterial06, target)
```

#### ManufacturingProcess32 and BiologicalMaterial06 are the top predictors. Both ManufacturingProcess and BiologicalMaterial contribute to the list if we consider the SVM algorithm.


